---
title: "Evo_Models"
author: "Agneesh Barua"
date: "05/08/2019"
output: html_document

---

```{r}
library(ape)
library(geiger)
library(phytools)
library(motmot)
library(tidyverse)
```

#MOTMOT tests for evolutionary models

##BPP.CNP
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(BPP.CNP) %>% as.matrix()
```

###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's Lambda
####Pagel’s lambda is a measure of phylogenetic ‘signal’ in which the degree to which shared history of taxa has driven trait distributions at tips. In this model, internal branch lengths are transformed by the lambda parameter value. When the parameter lambda equals 1, branches are transformed by multiplying by 1 and so the model is equal to Brownian motion (high phylogenetic signal). Values of lambda under 1 suggest there has been less influence of shared history on trait values at the tips. Finally, a lambda value of 0 indicates no phylogenetic influence on trait distributions, and is equivalent to a ‘star phylogeny’ with no shared branch lengths. (_taken from MOTMOT vignette_)
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

####We now compare the relative fit of the BM and lambda models. We can use the `stats` function `pchisq` to obtain a p value by testing using a chi-squared distribution
```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```
####Additionally there is a large small-sample Akaike Information Criterion (AICc) difference between the two models:
```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
####Kappa is used as a measure of punctuated evolution and spans values of 0-1 (Pagel 1997:1999). A Kappa value of 1 is equivalent to BM, and 0 indicates trait change occurs at events of speciation. Here there is evidence of punctuated evolution. transformPhylo.ML also allows users to see the the phylogeny transformed by model parameters. As an example, we show the original, BM model phylogeny and compare this with the phylogeny transformed by the Kappa phylogeny.  (_taken from MOTMOT vignette_)

```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```

###OU process
####The OU model allows for modelling of attraction to a optimum value, alpha (Hansen 1997; Butler and King 2004). This model again is similar to the Brownian motion model, but models the strength of attraction to alpha.
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```
###Delta
####Delta indicates a decrease or increase in the rate of trait evolution through time (Pagel 1997; 1999); a value of 1 is equivalent to Brownian motion, < 1 indicates a slow-down, and > 1 is indicates greater change closer to the present. Here we find a Maximum likelihood estimated for Delta of 2.23 but the CI spans < 1 to > 4, so it is not possible to conclusively support a change in the rate of evolution through time.

```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```


###ACDC model (Blomberg et al. 2003). 
####This model allows for exponential changes in the rate of evolution in the history of a clade.
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.


```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Kappa","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##CRISP
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(CRISP) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```
###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##CTL
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(CTL) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```
###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = T)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##GF
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(GF) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```
###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##KSPI
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(KSPI) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```
###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##LAAO
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(LAAO) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```
###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##SVMP
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/OUwie/SVMP_OUwie_ManyOptions_optima.csv", row.names = 1, header = T)

trait<- trait %>% select(SVMP_clr) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```

###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##SVSP
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(SVSP) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```
###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##TFTx
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(TFTx) %>% as.matrix()
```

###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```

###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##vPLA2
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(vPLA2) %>% as.matrix()
```


###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```

###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


##ePLA2
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)

trait<- trait %>% select(ePLA2) %>% as.matrix()
```

###Browninan (BM) null model
```{r}
bm.ml <- transformPhylo.ML(phy = tree, y = trait, model = "bm")
bm.ml
```

###Pagel's lamnda
```{r}
lambda.ml <- transformPhylo.ML(phy = tree, y = trait, model = "lambda",profilePlot = F)
lambda.ml
```

```{r}
p.value <- 1 - pchisq(lambda.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value
```

```{r}
bm.ml$AICc - lambda.ml$AICc
```

###Kappa
```{r}
kappa.ml <- transformPhylo.ML(phy =tree, y = trait, 
    model = "kappa", profilePlot = FALSE, returnPhy = TRUE)
par(mfrow = c(1, 2))
plot.phylo(tree, show.tip.label = FALSE, no.margin = TRUE)
mtext("Original phylogeny", 3, cex = 0.7, line = -1)
plot.phylo(kappa.ml$kappaPhy, show.tip.label = FALSE, no.margin = TRUE)
mtext("Kappa model phylogeny", 3, cex = 0.7, line = -1)
mtext("Kappa = 1e-8", 3, cex = 0.7, line = -2)
```

###OU
```{r}
ou.ml <- transformPhylo.ML(phy = tree, y = trait, model = "OU", profilePlot = F, upperBound = 2)

ou.ml
```

```{r}
p.value.OU <- 1 - pchisq(ou.ml$MaximumLikelihood - bm.ml$logLikelihood, 1)
p.value.OU   
```

```{r}
bm.ml$AICc - ou.ml$AICc
```

###Delta
```{r}
delta.ml <- transformPhylo.ML(tree, y = trait, 
    model = "delta")
delta.ml
```

###ACDC
```{r}
acdc.ml <- transformPhylo.ML(phy = tree, y = trait, 
                             model = "ACDC", profilePlot = F)

acdc.ml
```

```{r}
p.value.acdc <- 1 - pchisq(acdc.ml$MaximumLikelihood - bm.ml$logLikelihood, 
                        1)
p.value.acdc
```

####We constrain the ‘upperBound’ to < 0, this is equivalent to the Early Burst model (Harmon et al. 2010) fit in geiger.
```{r}
eb.ml<-transformPhylo.ML(phy = tree, y = trait, model = "ACDC", 
                  profilePlot = F, upperBound = -1e-06, print.warning = FALSE)
eb.ml
```
The estimate of -1e-6 for the exponential decrease parameter, which means the model is effectively equivalent to Brownian motion.

```{r}
aic.vals<-setNames(c(bm.ml$AICc,ou.ml$AICc,delta.ml$AICc,acdc.ml$AICc,eb.ml$AICc),c("BM","OU","Delta","ACDC","EB"))

aic.vals
```

```{r}
aic.w(aic.vals)
```


###############################################################################################################
#Evolution models using `fitContinuous` from `geiger` 
```{r}
tree<-read.newick("./Final_tree.tre")
trait<-read.csv("./data/scaleddata/scaled_data_fot_dtt.csv", row.names = 1, header = T)
```

##BPP.CNP
```{r}
dat<-setNames(trait$BPP.CNP,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##CRISP
```{r}
dat<-setNames(trait$CRISP,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##CTL
```{r}
dat<-setNames(trait$CTL,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##GF
```{r}
dat<-setNames(trait$GF,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##KSPI
```{r}
dat<-setNames(trait$KSPI,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##LAAO
```{r}
dat<-setNames(trait$LAAO,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##SVMP

```{r}
dat<-setNames(trait$SVMP,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##SVSP
```{r}
dat<-setNames(trait$SVSP,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```


##TFTx
```{r}
dat<-setNames(trait$TFTx,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```

##vPLA2
```{r}
dat<-setNames(trait$vPLA2,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```


##ePLA2
```{r}
dat<-setNames(trait$ePLA2,rownames(trait))
```

```{r, warning=FALSE}
fitBM<-fitContinuous(tree,dat)
fit_lambda <- fitContinuous(tree, dat, model="lambda")
fitOU<-fitContinuous(tree,dat,model="OU")
fitEB<-fitContinuous(tree,dat,model="EB")
```

```{r}
fitBM$opt
```

```{r}
fit_lambda$opt
```

```{r}
fitOU$opt
```

```{r}
fitEB$opt
```

```{r}
aic.vals<-setNames(c(fitBM$opt$aicc,fitOU$opt$aicc,fitEB$opt$aicc),
                   c("BM","OU","EB"))
aic.vals
```

```{r}
aic.w(aic.vals)
```